Computational methods in landscape ecology

### 2024-10-19 Sections/09-conclusion.qmd
Clean changes MJ

| Topic | Programming language | Software | Reference |
|:---:|:---:|:---:|:---:|
| Spatial patterns | R | sf | [@Pebesma2018SimpleFeatures] |
| Spatial patterns | R | terra | [@Hijmans2024TerraSpatial] |
| Spatial patterns | R | landscapemetrics | [@Hesselbarth2019LandscapemetricsOpensource] |
| Spatial patterns | R | multilandr | [@Huais2024MultilandrPackage] |
| Spatial patterns | R | bespatial | [@Nowosad2024BespatialBoltzmann] |
| Spatial patterns | R | rasterdiv | [@Rocchini2021RasterdivInformation] |
| Spatial patterns | R | LandComp | [@Konrad2023AdaptingMultiscale] |
| Spatial patterns | R | geodiv | [@Smith2021GeodivPackage] |
| Spatial patterns | R | glcm | [@Zvoleff2020GlcmCalculate] |
| Spatial patterns | R | vectormetrics | [@Matuszek2024VectormetricsLandscape] |
| Spatial patterns | R | motif | [@Nowosad2021MotifOpensource] |
| Spatial patterns | Python | LecoS | [@jung2016lecos] |
| Spatial patterns | Python | GeoPandas | [@Jordahl2020GeopandasPython] |
| Spatial patterns | Python | Rasterio | [@Gillies2013RasterioGeospatial] |
| Spatial patterns | Python | PyLandStats | [@Bosch2019PyLandStatsOpensource] |
| Spatial patterns | Julia | GeoStats | [@Hoffimann2018GeoStatsjlHighperformance] |
| Spatial patterns | Julia | GeoInterface | [@JuliaGeoandcontributors2024GeoInterfaceJulia] |
| Spatial patterns | Stand-alone software | FRAGSTATS | [@McGarigal2023FRAGSTATSV4] |
| Spatial patterns | Stand-alone software | GuidosToolbox | [@Vogt2017GuidosToolboxUniversal] |
| Connectivity | R | ResistanceGA | [@Peterman2018ResistanceGAPackage] |
| Connectivity | R | lconnect | [@Mestre2023LconnectPackage] |
| Connectivity | R | grainscape | [@Chubaty2020ToolboxGrainscape] |
| Connectivity | Julia | Circuitscape | [@Anantharaman2019CircuitscapeJulia] |
| Connectivity | Julia | ConScape | [@vanMoorter2023AcceleratingAdvances] |
| Connectivity | Julia | Omniscape | [@Landau2021OmniscapejlSoftware] |
| Connectivity | Stand-alone software | Graphab | [@Foltete2012SoftwareTool] |
| Connectivity | Stand-alone software | Conefor | [@Saura2009ConeforSensinode] |
| Connectivity | Stand-alone software | LSCorridors | [@Ribeiro2017LandScapeCorridors] |
| Sampling | Python | ZonalMetrics | [@Adamczyk2017ZonalMetricsPython] |
| Simulations | R | NLMR | [@Sciaini2018NLMRLandscapetools] |
| Simulations | R | rflsgen | [@Justeau-Allaire2022FragmentedLandscape] |
| Simulations | Python | NLMpy | [@Etherington2015NLMpyPython] |
| Simulations | Python | Pathwalker | [@UnnithanKumar2022PathwalkerNew] |
| Simulations | Julia | NeutralLandscapes | [@Poisot2023NeutralLandscapesJulia] |
| Simulations | NetLogo | EFForTS-LGraf | [@Salecker2019EFForTSLGrafLandscape] |
| Simulations | Stand-alone software | GradientLand | [@Cambui2015GradientLandSoftware] |
| Simulations | Stand-alone software | Landscape Generator | [@vanStrien2016ImprovedNeutral] |
| Simulations | Stand-alone software | HexSim | [@Schumaker2018HexSimModeling] |
| Simulations | Stand-alone software | RangeShifter2.0 | [@Bocedi2021RangeShifter20] |
| Genetics | R | graph4lg | [@Savary2021Graph4lgPackage] |
| Genetics | R | landgenreport | [@Gruber2015LandgenreportNew] |
| Genetics | R | adegenet | [@Jombart2008AdegenetPackage] |
| Genetics | R | ape | [@Paradis2019Ape50] |
| Genetics | Python | Python based scripts | [@Etherington2011PythonBased] |
| Genetics | Python | SDMtoolbox 2.0 | [@Brown2017SDMtoolbox20] |
| Planning | R | prioritizr | [@Hanson2024PrioritizrSystematic] |
| Planning | R | CoCo | [@vanMantgem2023CocoConservation] |
| Planning | R | restoptr | [@Justeau-Allaire2023RestoptrPackage] |
| Planning | Python | CAPTAIN | [@Silvestro2022ImprovingBiodiversity] |

@@

Alongside the previously introduced methodological advances specific to landscape ecology, the research field will also benefit from technological innovations from related disciplines. This includes advances in remote sensing technology with increasing data availability and diversity from a range of sources, including satellites, unmanned aerial vehicles, and ground-based sensor systems.
These systems provide both passive sensors, such as multispectral, hyperspectral, and thermal sensors and active optical sensors, such as LiDAR and SAR [@Crowley2020RemoteSensings].
Related to this, there is a growing number of up- and downscaling approaches for remote sensing data which makes a wider range of products accessible to scientists and spatial planners [@Markham2023ReviewMethods].
In combination with other increasingly available data sources, e.g, citizen science data, ecology has entered the era of big data, requiring methods that can handle heterogeneous data and high-throughput computing resources [@Farley2018SituatingEcology].
Due to its high flexibility and performance artificial intelligence has become popular in ecology, and deep learning and machine learning algorithms can be used for, e.g., mapping, classifying and extracting features, modeling, or predicting [@Christin2019ApplicationsDeep; @Stupariu2022MachineLearning; @Pichler2023MachineLearning].
Furthermore, quantum computing may additionally offer a potential pathway to surpass the limits of current computational technologies [@Woolnough2023QuantumComputing].
Last, landscape ecology will, as many other research fields, benefit from a general paradigm shift related to data sharing, management, and documentation [@Tenopir2020DataSharing].

>>>>

| Topic | Programming language | Software | Reference |
|:---:|:---:|:---:|:---:|
| Spatial patterns | R | sf | [@Pebesma2018SimpleFeatures] |
| Spatial patterns | R | terra | [@Hijmans2024TerraSpatial] |
| Spatial patterns | R | landscapemetrics | [@Hesselbarth2019LandscapemetricsOpensource] |
| Spatial patterns | R | multilandr | [@Huais2024MultilandrPackage] |
| Spatial patterns | R | bespatial | [@Nowosad2024BespatialBoltzmann] |
| Spatial patterns | R | rasterdiv | [@Rocchini2021RasterdivInformation] |
| Spatial patterns | R | LandComp | [@Konrad2023AdaptingMultiscale] |
| Spatial patterns | R | geodiv | [@Smith2021GeodivPackage] |
| Spatial patterns | R | glcm | [@Zvoleff2020GlcmCalculate] |
| Spatial patterns | R | vectormetrics | [@Matuszek2024VectormetricsLandscape] |
| Spatial patterns | R | motif | [@Nowosad2021MotifOpensource] |
| Spatial patterns | Python | GeoPandas | [@Jordahl2020GeopandasPython] |
| Spatial patterns | Python | Rasterio | [@Gillies2013RasterioGeospatial] |
| Spatial patterns | Python | PyLandStats | [@Bosch2019PyLandStatsOpensource] |
| Spatial patterns | Python | LecoS | [@Jung2016LecoSPython] |
| Spatial patterns | Julia | GeoStats | [@Hoffimann2018GeoStatsjlHighperformance] |
| Spatial patterns | Julia | GeoInterface | [@JuliaGeoandcontributors2024GeoInterfaceJulia] |
| Spatial patterns | Stand-alone software | FRAGSTATS | [@McGarigal2023FRAGSTATSV4] |
| Spatial patterns | Stand-alone software | GuidosToolbox | [@Vogt2017GuidosToolboxUniversal] |
| Connectivity | R | ResistanceGA | [@Peterman2018ResistanceGAPackage] |
| Connectivity | R | lconnect | [@Mestre2023LconnectPackage] |
| Connectivity | R | grainscape | [@Chubaty2020ToolboxGrainscape] |
| Connectivity | Julia | Circuitscape | [@Anantharaman2019CircuitscapeJulia] |
| Connectivity | Julia | ConScape | [@vanMoorter2023AcceleratingAdvances] |
| Connectivity | Julia | Omniscape | [@Landau2021OmniscapejlSoftware] |
| Connectivity | Stand-alone software | Graphab | [@Foltete2012SoftwareTool] |
| Connectivity | Stand-alone software | Conefor | [@Saura2009ConeforSensinode] |
| Connectivity | Stand-alone software | LSCorridors | [@Ribeiro2017LandScapeCorridors] |
| Sampling | Python | ZonalMetrics | [@Adamczyk2017ZonalMetricsPython] |
| Simulations | R | NLMR | [@Sciaini2018NLMRLandscapetools] |
| Simulations | R | rflsgen | [@Justeau-Allaire2022FragmentedLandscape] |
| Simulations | Python | NLMpy | [@Etherington2015NLMpyPython] |
| Simulations | Python | Pathwalker | [@UnnithanKumar2022PathwalkerNew] |
| Simulations | Julia | NeutralLandscapes | [@Poisot2023NeutralLandscapesJulia] |
| Simulations | NetLogo | EFForTS-LGraf | [@Salecker2019EFForTSLGrafLandscape] |
| Simulations | Stand-alone software | GradientLand | [@Cambui2015GradientLandSoftware] |
| Simulations | Stand-alone software | Landscape Generator | [@vanStrien2016ImprovedNeutral] |
| Simulations | Stand-alone software | HexSim | [@Schumaker2018HexSimModeling] |
| Simulations | Stand-alone software | RangeShifter2.0 | [@Bocedi2021RangeShifter20] |
| Genetics | R | graph4lg | [@Savary2021Graph4lgPackage] |
| Genetics | R | landgenreport | [@Gruber2015LandgenreportNew] |
| Genetics | R | adegenet | [@Jombart2008AdegenetPackage] |
| Genetics | R | ape | [@Paradis2019Ape50] |
| Genetics | Python | Python based scripts | [@Etherington2011PythonBased] |
| Genetics | Python | SDMtoolbox 2.0 | [@Brown2017SDMtoolbox20] |
| Planning | R | prioritizr | [@Hanson2024PrioritizrSystematic] |
| Planning | R | CoCo | [@vanMantgem2023CocoConservation] |
| Planning | R | restoptr | [@Justeau-Allaire2023RestoptrPackage] |
| Planning | Python | CAPTAIN | [@Silvestro2022ImprovingBiodiversity] |

@@

Alongside the previously introduced methodological advances specific to landscape ecology, the research field will also benefit from technological innovations from related disciplines. 
This includes advances in remote sensing technology with increasing data availability and diversity from a range of sources, including satellites, unmanned aerial vehicles, and ground-based sensor systems.
These systems provide both passive sensors, such as multispectral, hyperspectral, and thermal sensors and active optical sensors, such as LiDAR and SAR [@Crowley2020RemoteSensings].
Related to this, there is a growing number of up- and downscaling approaches for remote sensing data which makes a wider range of products accessible [@Markham2023ReviewMethods].
In combination with other increasingly available data sources, e.g, citizen science data, ecology has entered the era of big data, requiring methods that can handle heterogeneous data and high-throughput computing resources [@Farley2018SituatingEcology].
Due to its high flexibility and performance artificial intelligence has become popular in ecology, and deep learning and machine learning algorithms can be used for, e.g., mapping, classifying and extracting features, modeling, or predicting [@Christin2019ApplicationsDeep; @Stupariu2022MachineLearning; @Pichler2023MachineLearning].
Furthermore, quantum computing may additionally offer a potential pathway to surpass the limits of current computational technologies [@Woolnough2023QuantumComputing].
Last, landscape ecology will, as many other research fields, benefit from a general paradigm shift related to data sharing, management, and documentation [@Tenopir2020DataSharing].
